### 2024-07-27 mixed-models.qmd
20240727 - peak ages and age of zero expected fantasy points

Plots of players' raw fantasy points data by age are in Figures [-@fig-rawFantasyPointsByAgeQB]–[-@fig-rawFantasyPointsByAgeTE].

@@

@@

@@

@@

@@

pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$ageCentered20)

@@

@@

@@

@@

@@

player_stats_seasonal_offense_subsetCC$fantasyPoints_gam <- predict(
  object = pointsPerSeason_gam,
  newdata = player_stats_seasonal_offense_subsetCC
)
```

Plots of players' model-implied fantasy points by age are in Figures [-@fig-individualFantasyPointsByAgeQB]–[-@fig-individualFantasyPointsByAgeTE].

@@

@@

@@

@@

>>>>

#### Quarterbacks {#sec-fantasyPointsByAgeExperienceRawDataQB}

A plot of Quarterbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeQB.

@@

#### Fullbacks {#sec-fantasyPointsByAgeExperienceRawDataFB}

A plot of Fullbacks' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeFB.

@@

#### Running Backs {#sec-fantasyPointsByAgeExperienceRawDataRB}

A plot of Running Backs' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeRB.

@@

#### Wide Receivers {#sec-fantasyPointsByAgeExperienceRawDataWR}

A plot of Wide Receivers' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeWR.

@@

#### Tight Ends {#sec-fantasyPointsByAgeExperienceRawDataTE}

A plot of Tight Ends' raw fantasy points data by age is in @fig-rawFantasyPointsByAgeTE.

@@

pointsPerSeason_positionAge_newData$ageCentered20 <- pointsPerSeason_positionAge_newData$age - 20
pointsPerSeason_positionAge_newData$ageCentered20Quadratic <- pointsPerSeason_positionAge_newData$ageCentered20 ^ 2
pointsPerSeason_positionAge_newData$years_of_experience <- floor(pointsPerSeason_positionAge_newData$age - 22) # assuming that most players start at age 22 (i.e., rookie year) and thus have 1 year of experience at age 23
pointsPerSeason_positionAge_newData$years_of_experience[which(pointsPerSeason_positionAge_newData$years_of_experience < 0)] <- 0

@@

#### Quadratic Model {#sec-fantasyPointsByAgePlotsQuadratic}

@@

#### Quadratic Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsQuadraticDepth}

@@

#### Generalized Additive Model {#sec-fantasyPointsByAgePlotsGAM}

@@

#### Generalized Additive Model: Top of Depth Chart {#sec-fantasyPointsByAgePlotsGAMdepth}

@@

player_stats_seasonal_offense_subsetCC$fantasyPoints_gam <- predict(
  object = pointsPerSeason_gam,
  newdata = player_stats_seasonal_offense_subsetCC
)

zeroAge <- pointsPerSeason_positionAge_newData %>% 
  group_by(positionFactor) %>% 
  filter(fantasyPoints_gam < 0) %>% 
  slice(which.min(age))

peakAge <- pointsPerSeason_positionAge_newData %>% 
  group_by(positionFactor) %>% 
  slice(which.max(fantasyPoints_gam))

peakAge2 <- pointsPerSeason_positionAge_newData %>% 
  filter(age > 22) %>% 
  group_by(positionFactor) %>% 
  slice(which.max(fantasyPoints_gam))

qbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "QB")], 0)
fbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "FB")], 0)
rbPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "RB")], 0)
wrPeakAge <- round(peakAge$age[which(peakAge$positionFactor == "WR")], 0)
wrPeakAge2 <- round(peakAge2$age[which(peakAge$positionFactor == "WR")], 0)
tePeakAge <- round(peakAge$age[which(peakAge$positionFactor == "TE")], 0)

qbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "QB")], 0)
fbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "FB")], 0)
rbZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "RB")], 0)
wrZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "WR")], 0)
teZeroAge <- round(zeroAge$age[which(zeroAge$positionFactor == "TE")], 0)
```

#### Quarterbacks {#sec-fantasyPointsIndividualByAgePlotsQB}

A plot of Quarterbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeQB.
The model-implied peak of Quarterbacks' fantasy points is at age `{r} qbPeakAge`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} qbZeroAge`.

@@

#### Fullbacks {#sec-fantasyPointsIndividualByAgePlotsFB}

A plot of Fullbacks' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeFB.
The model-implied peak of Fullbacks' fantasy points is at age `{r} fbPeakAge`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} fbZeroAge`.

@@

#### Running Backs {#sec-fantasyPointsIndividualByAgePlotsRB}

A plot of Running Backs' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeRB.
The model-implied peak of Running Backs' fantasy points is at age `{r} rbPeakAge`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} rbZeroAge`.

@@

#### Wide Receivers {#sec-fantasyPointsIndividualByAgePlotsWR}

A plot of Wide Receivers' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeWR.
The model-implied peaks of Running Backs' fantasy points are at ages `{r} wrPeakAge` and `{r} wrPeakAge2`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} wrZeroAge`.

@@

#### Tight Ends {#sec-fantasyPointsIndividualByAgePlotsTE}

A plot of Tight Ends' model-implied fantasy points by age is in @fig-individualFantasyPointsByAgeTE.
The model-implied peak of Running Backs' fantasy points is at age `{r} tePeakAge`.
The model-predicted value of zero fantasy points for Quarterbacks is at `{r} teZeroAge`.
